SDSS-IV MaNGA: Radial Profiles of Specific Star Formation Rate in Nearby Interacting Galaxies

### 2020-06-17 maintex.tex
added SDSS acknowledgement

\begin{abstract}
We study the spatial profile of the star formation rate (SFR) in a sample of 202 interacting galaxies within the MaNGA (Mapping Nearby Galaxies at Apache Point Observatory) survey. We find that the specific star formation rate (sSFR) is centrally enhanced in the interacting galaxies by $\sim$0.25 $\pm$ 0.1 but that there is no enhancement or suppression in their disks. We further investigate how these profiles behave as a function of stellar mass, mass ratio, and projected separation. We find that the sSFR profiles are unaffected by the stellar masses of the interacting galaxies; however, they are affected by the mass ratio between the galaxies where the galaxies with the smallest mass ratio feature the largest enhancement. We also find that, in pairs with large mass ratios ($|\Delta$log(M)$|$ $=$ 0.5 $-$ 1.5), the less massive galaxy of the pair has slightly higher levels of sSFR enhancement, $\sim$0.1 dex. We also find that the level of sSFR enhancement increases with closer projected separations; however, the closest pairs, $\le$5 kpc, show no signs of sSFR enhancement/suppression. We suspect that these close pairs are merging galaxies which have passed their first pericenter, but the merger induced gas-inflows have yet to reach the centers of the galaxies. We also set a simple model for the future mass growth of the galaxies due to the merger induced star formation. We predict that the paired galaxies in our sample will see a mass growth rate which is XX$\times$ higher than secular mass growth. 
\end{abstract}

@@

%%%Galaxy Evolution%%%
In the $\Lambda$CDM model, galaxy evolution is a hierarchical process. Over the lifetime of the universe, smaller galaxies have iteratively merged forming more massive galaxies. In fact, cosmological hydrodynamical simulations have shown that repeated merger events may be responsible for as much as $\sim$60\% of stellar mass in massive galaxies like M87 (e.g. \citet{Rodriguez-Gomez:2016}; \citet{Pillepich:2018}).

@@

%%%Observations%%%
Centrally enhanced star formation has also been observed in large spectroscopic surveys including; CfA2  \citep{Barton:2000, Woods:2006}, 2dF \citep{Lambas:2003}, AEGIS \citep{Lin:2007}, SDSS \citep{Ellison:2008}, COSMOS \citep{Kartaltepe:2007,Xu:2012}, PRIMUS \citep{Wong:2011}, and GAMA \citep{Robotham:2014}.

This star formation enhancement increases with closer projected separations between the two galaxies \citep{Ellison:2008, Scudder:2012} but has still shown to be detectable at projected separations of 150 kpc \citep{Patton:2013}. The level of the star formation enhancement also increases with smaller stellar mass ratios, $|\Delta log {\it M} |$, between the two galaxies \citep{Ellison:2008}.

%%%Integral Field Spectroscopy%%%
With the recent large integral field spectroscopic (IFS) surveys, interacting galaxies can be studied with unprecedented spatial detail. Where previous spectroscopic surveys were limited to the nucleus of the interacting galaxies, the new IFS surveys will be able to study both the nuclei and the disks of the interacting galaxies.

The IFS surveys, CALIFA\footnote{Calar Alto Legacy Integral Field Area} and MaNGA\footnote{Mapping Nearby Galaxies at Apache Point Observatory}, have already been used to study the SFR of interacting galaxies in merging galaxies. The SFR was found to be centrally enhanced by a factor of 2$-$3$\times$ \citep{Barrera-Ballesteros:2015, Pan:2019}, which is in agreement with previous single-fiber spectroscopic surveys. These surveys also studied the SFR in the disk of the paired galaxies; however, there is a discrepancy between the results. \citet{Barrera-Ballesteros:2015} found a suppression to the SFR of $\sim$0.74$\times$ while \citet{Pan:2019} found an enhancement of 1.22$-$1.58$\times$.

The MaNGA survey has also been used to study the radial profiles of SFR in post merger galaxies, which are the galaxies which have recently coalesced. The post merger galaxies showed a central enhancement to the SFR surface density of 2.5$\times$ and an enhancement in their disks of 1.26$\times$ \citep{Thorp:2019}.

%%%Purpose%%%
In our previous work, \citet{Fu:2018}, we built a sample of paired galaxies in the MaNGA survey where both components of the pair were contained within field of view of a single IFU. We found that approximately 5.7\% of the MaNGA galaxies have a companion galaxies contained with the IFU. In this work, we will use a version of this sample updated for the recent MaNGA Data Release 16 (DR16) and further supplement this with a sample of companion galaxies identified outside the field of view of the IFUs.

In this work we will expand upon the previous works by studying how the spatial profile of SFR behaves as a function of the stellar mass of the paired galaxy, the mass ratio between the galaxy pairs, and the projected separation between the galaxies. We will also model how the merger-induced star formation will affect the future mass growth of the interacting galaxies.

% Organization
This paper is organized as follows; in Section \ref{sec:data} we will discuss the detail of the MaNGA survey, in Section \ref{sec:analysis} we will discuss how we extract emission lines from the MaNGA survey and how we calculate the star formation from them, in Section \ref{sec:pair} we discuss the construction of our pair sample, in Section \ref{sec:control} we discuss the construction of our control sample, in Section \ref{sec:radial} we build the radial profiles for the pair and control galaxies, in Section \ref{sec:results} we discuss the results of the survey, in Section \ref{sec:discussion} we discuss our results in the context of other surveys, and in Section \ref{sec:sum} we summarize and conclude the work.

@@

MaNGA is an a IFS survey based at Apache Point Observatory (APO) which used the SDSS 2.5 meter telescope along with two dual-channel BOSS spectrographs. MaNGA captures spectra through 17 integral field units (IFUs) with variable numbers of fibers; 19, 37, 61, 91, and 127 fibers covering 12.5\arcsec, 17.5\arcsec, 22.5\arcsec, 27.5\arcsec, and 32.5\arcsec on the sky respectively. MaNGA is an optical survey with a spectral coverage of 3600$-$10,300 \AA with a resolution of R $\sim$ 2000 and a median spectra coverage of 2.5\arcsec FWHM \citep{Bundy:2015}.

MaNGA targets galaxies from a subset of 41,154 galaxies with from the NASA-Sloan Atlas (NSA) with a redshift range of 0.01 $<$ z $<$ 0.15 and a luminosity range of -17.7 $<$ $\mathcal{M_{\rm i}}$ $<$ -24.0, where $\mathcal{M_{\rm i}}$ is the rest frame i-band magnitude within the NSA survey's elliptical Petrosian apertures. MaNGA plans to cover 10,000 galaxies with a flat stellar mass distribution at two spatial coverages, 1.5 \reff\ and 2.5 \reff\ (where \reff\ is the radius which contains 50\% of the galaxy's total light). In this work we use the data from the internal data release, MPL-8, which covers 6142 unique galaxies.

@@

Because of the survey's flat stellar mass distribution and uniform spatial coverage, these samples have a unique ``banana" shaped distribution in redshift and luminosity, see Figure \ref{fig:Mi-z}. Because the Secondary and Color-Enhanced samples use a wider spatial coverage than the Primary sample, the Secondary and Color-Enhanced Samples will posses galaxies of higher redshifts than galaxies at the same stellar masses in the Primary sample.

%%% NSA and Simard+11
We further use photometric data from the NSA catalog and \citet{Simard:2011} (from now on \simard). We use the r-band elliptical Petrosian apertures from the NSA catalog and the r-band elliptical S\'ersic apertures from \simard\ to quantify the geometry of the outside-IFU and inside-IFU samples respectively. We further use the total stellar masses extracted from these apertures to quantify the total stellar masses of the galaxies, the NSA catalog contains their own calculation for stellar mass and we use the stellar mass calculation from \citet{Mendel:2014} for the \simard\ apertures.

@@

%%% SPFIT
We use our own {\sc IDL} based spectral fitting code, {\sc SPFIT}, to model the spectra from the MaNGA datacubes. {\sc SPFIT} simultaneously fits emission lines and the stellar continuum with the Levenberg-Marquardt nonlinear least-squares minimization algorithm \citep{Fu:2018}. Emission line are parameterized as a Gauss-Hermite series and the stellar continuum is the sum of simple stellar populations (SSPs) convolved with the line-of-sight velocity distribution (LOSVD). The algorithm provides us with measurements of emission line fluxes, equivalent widths, velocities, and velocity dispersions of 19 emission lines and measurements of stellar masses, ages, [Fe/H], and kinematics of the stellar populations.

%%% Reddening
We use the extinction curve from \citet{Cardelli:1989} with updated coefficients from \citet{ODonnell:1994}. The extinction in parameterized as R$_V$ $\equiv$ A$_V$/E(B$-$V) $=$ 3.1 where we estimate the value of the V-band extinction, A$_V$, by comparing the H$\alpha$/H$\beta$ to the expected value of 2.85 for case-B recombination. We apply the extinction correction to all of the extracted emission lines.

\subsection{Star Formation Rates}
We classify galaxies as star forming using both the color-magnitude diagram (CMD) and the BPT diagram \citep{Baldwin:1981}. We show the CMD in Figure \ref{fig:cmd} along with demarcation lines which separate the blue cloud, red sequence, and green valley. We established the demarcation lines by collapsing the CMD to a color histogram for each of the three regions and then varied the slopes between the regions until we found the slopes which best fit the data. These demarcation lines are;

@@

Where NUV$-$r is the color and M$_i$ is the i-band magnitude.

The BPT diagnostic uses emission line ratios to separate ionization sources, star formation and active galactic nuclei (AGN). The two ionization sources can be separated by either the theoretical maximum starburst line from \citet{Kewley:2001} or the analytically fitted line from \citet{Kauffmann:2003}, given below in Equations \ref{eq:kewley} and \ref{eq:kauffmann} respectively.

@@

Where \OIII$\lambda$5007, H$\beta$, \NII$\lambda$6584, and H$\alpha$ are emission line fluxes extracted from the spectra within the inner 2 arcseconds of the galaxies. It has also been shown that hot low mass evolved stars (HOLMES) in retired galaxies can mimic HII regions and AGN on the BPT diagram \citep{Stasinska:2008}. We apply an H$\alpha$ equivalent width (EW) cut of \ewha\ $>$ 3\angstrom to remove these retired galaxies \citet{Cid-Fernandes:2011}. In this work, we define star forming galaxies as those which are on the blue cloud in the CMD and which is below the \citet{Kewley:2001} on the BPT diagram.

We measure the SFR from the H$\alpha$ luminosity, L$_{H\alpha}$. The H$\alpha$ luminosity is calculated from the flux of the H$\alpha$ emission line. We use the SFR formula, Equation \ref{eq:sfr}, from \citet{Murphy:2011} which uses a Kroupa IMF, Solar metallicity, a constant SFR at an age of 100 Myr, and Case-B recombination.

@@

We normalize the SFR by the local stellar mass (the stellar mass contained with a given spaxel), calculated from SPFIT, giving us the specific star formation rate (sSFR).

\begin{equation}
sSFR = \frac{SFR}{M^*}
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Pair Sample}\label{sec:pair}

@@

We use SDSS photometric objects to create a catalog of objects within the fields of view of the MaNGA IFUs. The extract their spectra from the MaNGA datacubes through a 1\arcsec aperture and fit the spectra with {\sc SPFIT}. The spectra is then sorted into different categories; ``good" galaxy spectra, broad-line AGN, foreground star, foreground/background galaxies, or poor S/N objects. The ``good" galaxy spectra are the objects whose spectra is well modeled by {\sc SPFIT} at the target galaxy's redshift, whether it is the target galaxies itself or a nearby companion galaxy. This means that the companion galaxy can be within $\pm$3000 km s$^{-1}$ of the MaNGA target. We found 6573 ``good" objects, 57 broad-line AGN, 836 foreground stars, 319 foreground/background galaxies, and 1546 objects with poor S/N.

From the 6573 galaxies with good spectra, 404 of the MaNGA IFUs have multiple objects within the IFU and within a relative velocity of $\Delta$v $<$ 500 km s$^{-1}$. 124 of the galaxies are star forming galaxies (SFGs). Finally, 41 SFGs are covered by \simard\ and \citet{Mendel:2014}.

Some of the galaxies in our sample have multiple companions. From the whole set of 404 IFUs, there are 327 pairs, 67 triplets, 7 quadruplets, and 1 quintuplet.

We characterize the geometry of the inside-IFU galaxies with the r-band elliptical S\'ersic apertures and we characterize the total stellar mass of the galaxies with the stellar masses from \citet{Mendel:2014}

@@

The second pairs sample is the outside-IFU pairs sample. This sample includes MaNGA galaxies which have companion galaxies found outside the field of view of its IFU. We identify these external pair from the NSA catalog. We select galaxies which are within a projected separation, r$_p$ $<$ 50 kpc, and within a relative velocity, $\Delta$v $<$ 500 km s$^{-1}$.

The NSA catalog contains 641,409 galaxies from which we find 509 galaxies which are paired to MaNGA targets. MaNGA targets which have both an inside-IFU and an outside-IFU pair are left to the inside-IFU sample. After restricting the sample to SFGs, we have 161 MaNGA targets with paired galaxies outside of the IFU.

We use the r-band elliptical Petrosian apertures from the NSA catalog to describe the geometry of the galaxies in the outside-IFU sample. We also use the stellar masses from the NSA catalog for the total stellar mass of the galaxies.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Control Sample}\label{sec:control}
We build a sample of isolated control galaxies from the remaining MaNGA galaxies. Selecting the MaNGA galaxies which have no spectroscopic companions within r$_p$ $<$ 50 kpc and $\Delta$v $<$ 500 km s$^{-1}$ either inside or outside of the IFU. This gives us a set of 1862 SFGs and 1598 SFGs which are covered by both \simard\ and \citet{Mendel:2014}.

The SFR in galaxies has been shown to be dependent on both stellar mass and redshift \citep{Noeske:2007}. To account for this we will need to match pairs with similar mass and redshift controls. To do this we build a subset of 20 control galaxies for each paired galaxy.

First we select every control galaxy which is within a stellar mass of 0.1 dex and within in same MaNGA subsample, i.e. the Primary or Secondary sample. Knowing that MaNGA has a tight ``banana" shaped distribution in stellar mass and redshift, imposing the stellar mass cut and the MaNGA subsample cut will effectively set a $\sim$0.025 dex redshift cut.

We then downselect the control subset down to 20 controls using a random number generator so that each paired galaxies has the same number of selected controls. In the few cases where a paired galaxy fails to find at least 20 control galaxies, we iteratively expand the stellar mass limits by 0.1 dex until a set of 20 control galaxies can be made. A small set of paired required extra iterations to find 20 controls; 8 paired galaxies need an extra iteration, 1 paired galaxy need two extra iterations, and 1 paired galaxy needed three iterations. \\

The control galaxies will be treated using both the \simard\ aperture and \citet{Mendel:2014} masses when comparing them to the inside-IFU sample and the NSA apertures and masses when comparing them to the outside-IFU sample.

@@

As mentioned in Sections \ref{sec:pair} and \ref{sec:control} we use the r-band elliptical Petrosian apertures from the NSA catalog and the r-band S\'ersic apertures from \simard to define the geometries of the galaxies. We calculate the inclination angle, {\it i}, of the galaxies using the major-to-minor axis ratios from the elliptical apertures;

@@

The inclination angle is used to deproject the geometries of the galaxies. We use our catalog of photometric objects and the \simard\ catalog to mask foregournd/background objects. Objects in our photometric catalog are masked out with a 2\arcsec\ circular aperture and the galaxies from \simard\ are masked out with a 2.0 \reff\ elliptical aperture, using the r-band S\'ersic apertures from the catalog.

>>>>

\begin{abstract}
We study the spatial profile of the star formation rate (SFR) in a sample of 202 interacting galaxies within the MaNGA (Mapping Nearby Galaxies at Apache Point Observatory) survey. We find that the specific star formation rate (sSFR) is centrally enhanced in interacting galaxies by $\sim$0.25 $\pm$ 0.1 but that there is no enhancement or suppression in their disks. We further investigate how these profiles behave as a function of their stellar mass, mass ratio, and projected separation. We find that the sSFR profiles are unaffected by the stellar masses of the interacting galaxies; however, they are affected by the mass ratio between the galaxies where the galaxies with the smallest mass ratios feature the largest enhancement to their sSFR. We also find that, in pairs with large mass ratios ($|\Delta$log(M)$|$ $=$ 0.5 $-$ 1.5), the less massive galaxy of the pair has slightly higher levels of sSFR enhancement, $\sim$0.1 dex, than the more massive galaxies in the pair. We also find that the level of sSFR enhancement increases with closer projected separations; however, the closest pairs, $\le$5 kpc, show no signs of sSFR enhancement/suppression. We suspect that these close pairs are merging galaxies which have passed their first pericenter, but the merger induced gas-inflows have yet to reach the centers of these galaxies. We make a simple model for the future mass growth of the galaxies due to the merger induced star formation. We predict that the paired galaxies in our sample will see a mass growth rate in their centers which is 1.6$-$2.3$\times$ higher than secular mass growth, showing that the merger induced star formation can significantly alter the stellar mass distribution. 
\end{abstract}

@@

%%%Galaxy Evolution%%%
In the $\Lambda$CDM model, galaxy evolution is a hierarchical process. In this model, massive galaxies are the product of several past merger events of smaller galaxies. In fact, cosmological hydrodynamical simulations have shown that repeated merger events may be responsible for as much as $\sim$60\% of stellar mass in massive galaxies like M87 (e.g. \citet{Rodriguez-Gomez:2016}; \citet{Pillepich:2018}).

@@

%%%Observations%%%
Centrally enhanced star formation has also been observed in galaxy pairs with large spectroscopic surveys including; CfA2  \citep{Barton:2000, Woods:2006}, 2dF \citep{Lambas:2003}, AEGIS \citep{Lin:2007}, SDSS \citep{Ellison:2008}, COSMOS \citep{Kartaltepe:2007,Xu:2012}, PRIMUS \citep{Wong:2011}, and GAMA \citep{Robotham:2014}.

This star formation enhancement increases with closer projected separations between the two galaxies \citep{Ellison:2008, Scudder:2012} but has still shown to be detectable out to projected separations of 150 kpc \citep{Patton:2013}. The level of the star formation enhancement also increases with smaller stellar mass ratios, $|\Delta {\rm log} {\it M} |$, between the two galaxies \citep{Ellison:2008}. Similar mass galaxies show the strongest sSFR effects while large central galaxies interacting with small satellite galaxies show weaker sSFR effects. Because of this, most surveys will split their sample into major mergers, $|\Delta$log($M$)$|$ $\le$ 0.5, and minor mergers, 0.5 $\le$ $|\Delta$log($M$)$|$ $\le$ 1.0, where major mergers are expected to show the strongest interaction induced effects.

%%%Integral Field Spectroscopy%%%
With the recent large integral field spectroscopic (IFS) surveys, interacting galaxies can be studied with unprecedented spatial detail. These surveys allow astronomers to study the centers of merging galaxies more rigorously since apertures can be set to the physical scale of the galaxies instead of being bound by a fixed sky aperture. These IFS surveys will also allow us to see the extent of the centrally induced star formation and to see how the star formation in the disks of the galaxies are affected.

The IFS surveys, CALIFA\footnote{Calar Alto Legacy Integral Field Area} and MaNGA\footnote{Mapping Nearby Galaxies at Apache Point Observatory}, have already been used to study the SFR of interacting galaxies in merging galaxies. The SFR was found to be centrally enhanced by a factor of 2$-$3$\times$ \citep{Barrera-Ballesteros:2015, Pan:2019}, which is in agreement with the previous single-fiber spectroscopic surveys. These surveys also studied the SFR in the disk of the paired galaxies; however, there is a standing discrepancy between the different works. \citet{Barrera-Ballesteros:2015} found a suppression to the SFR of $\sim$0.74$\times$ while \citet{Pan:2019} found an enhancement of 1.22$-$1.58$\times$.

The MaNGA survey has also been used to study the radial profiles of SFR in post-merger galaxies, which are the galaxies which have recently coalesced. Hydrodynamical simulations predict a second burst of star formation as the two galaxies coalesce. This effect has been observed in a sample of post-merger galaxies which showed a central enhancement to the SFR surface density of 2.5$\times$ and an enhancement in their disks of 1.26$\times$ \citep{Thorp:2019}.

%%%Purpose%%%
In our previous work, \citet{Fu:2018}, we built a sample of paired galaxies in the MaNGA survey where both components of the pair were contained within field of view of a single integral field unit (IFU). We found that approximately 5.7\% of the MaNGA galaxies have a companion galaxy contained within the field-of-view of a single IFU. In this work, we will use a version of this sample updated for the recent MaNGA Data Release 16 (DR16) and we will further supplement this with a sample of companion galaxies identified outside the field-of-view of the MaNGA IFUs.

In this work we will study how the radial profiles of star formation in our pair sample is effected by tidal interactions. We will expand upon the previous studies of star formation in merging galaxies by studying how the spatial profile of the SFR behaves as a function of the stellar mass of the paired galaxy, the mass ratio between the galaxy pairs, and the projected separation between the galaxies.

Changes in the star formation rate will also affect the future mass growth in the interacting galaxies. We will model the fractional mass growth rate of the galaxies in our sample to see how the merger event influences the mass evolution of the paired galaxies. Since merging galaxies are already known to have centrally enhanced star formation rates, we would expect this to produce higher concentrations of stellar mass in their centers.

% Organization
This paper is organized as follows; in Section \ref{sec:data} we will discuss the properties of the MaNGA survey along with the photometric catalogs which we use characterize the stellar masses and geometries or our galaxies, in Section \ref{sec:analysis} we will discuss how we extract emission lines from the MaNGA survey and how we calculate star formation rates from them, in Section \ref{sec:pair} we discuss the construction of our pair samples, in Section \ref{sec:control} we discuss the construction of a sample of isolated, control galaxies, in Section \ref{sec:radial} we build the radial profiles for both the paired and control galaxies, in Section \ref{sec:results} we discuss the results of the constructed radial profiles, and in Section \ref{sec:sum} we discuss and summarize the finding of the work.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=\linewidth]{fig/Mi-z.pdf}
\caption[]{\textbf{Left:} The luminosity $-$ redshift distribution of the three MaNGA subsamples; The Primary Sample (blue), the Secondary Sample (red) and the Color-Enhanced Sample (green). \textbf{Right:} The stellar mass $-$ redshift distribution of the three samples. }
\label{fig:Mi-z}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

MaNGA is an a IFS survey based at Apache Point Observatory (APO) which used the SDSS (Sloan Digital Sky Survey) 2.5-meter telescope along with two dual-channel BOSS spectrographs. MaNGA captures spectra through 17 integral field units (IFUs) with variable numbers of fibers; 19, 37, 61, 91, and 127 fibers covering 12.5\arcsec, 17.5\arcsec, 22.5\arcsec, 27.5\arcsec, and 32.5\arcsec on the sky respectively. MaNGA is an optical survey with a spectral coverage of 3600$-$10,300 \AA\ with a resolution of R $\sim$ 2000 and a median spectra coverage of 2.5\arcsec\ FWHM \citep{Bundy:2015}.

MaNGA targets galaxies from a subset of 41,154 galaxies from the NASA-Sloan Atlas (NSA) with a redshift range of 0.01 $<$ z $<$ 0.15 and a luminosity range of -17.7 $<$ $\mathcal{M_{\rm i}}$ $<$ -24.0, where $\mathcal{M_{\rm i}}$ is the rest frame i-band magnitude within the survey's elliptical Petrosian apertures. MaNGA plans to cover 10,000 galaxies with a flat stellar mass distribution at two spatial coverages, 1.5 \reff\ and 2.5 \reff\ (where \reff\ is the radius which contains 50\% of the galaxy's total light). In this work we use the data from the internal data release, MPL-8, which covers 6142 unique galaxies.

@@

Because of the survey's flat stellar mass distribution and uniform spatial coverage, these samples have a unique ``banana" shaped distribution in redshift and luminosity, seen in Figure \ref{fig:Mi-z}. Because the Secondary and Color-Enhanced samples use a wider spatial coverage than the Primary sample, the Secondary and Color-Enhanced Samples will posses galaxies of higher redshifts than galaxies at the same stellar masses in the Primary sample. This distribution will be important later when we match control galaxies to paired galaxies based on their stellar masses and redshifts.

%%% NSA and Simard+11
We use photometric data from the NSA catalog and \citet{Simard:2011} (from now on \simard). We use the r-band elliptical Petrosian apertures from the NSA catalog and the r-band elliptical S\'ersic apertures from \simard\ to characterize the geometry of galaxies with pairs identified outside of the IFU (the outside-IFU sample) and galaxies identified within the IFU (the inside-IFU sample) respectively. We further use the total stellar masses extracted from these apertures to quantify the total stellar masses of the galaxies, the NSA catalog contains its own calculation for stellar mass and we use the stellar mass calculation from \citet{Mendel:2014} for the \simard\ apertures.

We use two different photometric catalogs here because the NSA catalog covers the entirety of the MaNGA sample, but does not adequately fit the geometry of galaxies with close companions. In many of the cases where we have a MaNGA target with an inside-IFU pair, the NSA catalog's algorithm fails to model the geometry of the MANGA target as the algorithm fails to separately deblend the two galaxies. We found that the algorithm used in the \simard\ catalog models the galaxies with close companions much better; however, the catalog does not completely cover our sample. We use both of the catalogs as a compromise to maintain the sample size in the outside-IFU sample, where the NSA catalog models the paired galaxies' geometry well, and to obtain better geometry in our inside-IFU sample, which is poorly fitted by the NSA catalog's algorithm.

@@

%%% SPFIT
We use our own {\sc IDL} based spectral fitting code, {\sc SPFIT}, to model the spectra from the MaNGA datacubes. {\sc SPFIT} simultaneously fits emission lines and the stellar continuum with the Levenberg-Marquardt nonlinear least-squares minimization algorithm \citep{Fu:2018}. Emission lines are parameterized as a Gauss-Hermite series and the stellar continuum is the sum of simple stellar populations (SSPs) convolved with the line-of-sight velocity distribution (LOSVD). The algorithm provides us with measurements of emission line fluxes, equivalent widths, velocities, and velocity dispersions of 19 emission lines and measurements of stellar masses, ages, [Fe/H], and the kinematics of the stellar populations.

%%% Reddening
We correct the emission lines for reddening using the extinction curve from \citet{Cardelli:1989} with updated coefficients from \citet{ODonnell:1994}. The extinction in parameterized as R$_V$ $\equiv$ A$_V$/E(B$-$V) $=$ 3.1, where we estimate the value of the V-band extinction, A$_V$, by comparing the H$\alpha$/H$\beta$ ratio to the expected value of 2.85 for case-B recombination. We apply the extinction correction to all of the extracted emission lines.

\subsection{Star Formation Rates}
To ensure that the observed emission lines are created by HII regions, we classify galaxies as star forming using both the color-magnitude diagram (CMD) and the BPT diagram \citep{Baldwin:1981}. We show the CMD in Figure \ref{fig:cmd} along with demarcation lines which separate the blue cloud, red sequence, and green valley. We established the demarcation lines by collapsing the CMD to a color histogram for each of the three regions. We then varied the slopes between the regions until we found the slopes which best fit the data. These demarcation lines are;

@@

Where NUV$-$r is the color and M$_i$ is the i-band magnitude from the NSA catalog.

The BPT diagnostic uses emission line ratios to separate ionization sources; star formation and active galactic nuclei (AGN). The two ionization sources are often separated by either the theoretical maximum starburst line from \citet{Kewley:2001} or the analytically fitted line from \citet{Kauffmann:2003}, given below in Equations \ref{eq:kewley} and \ref{eq:kauffmann} respectively.

@@

Where \OIII$\lambda$5007, H$\beta$, \NII$\lambda$6584, and H$\alpha$ are emission line fluxes extracted from the MaNGA spectra within the inner 2 arcseconds of the galaxies. It has also been shown that hot low mass evolved stars (HOLMES) in retired galaxies can mimic HII regions and AGN on the BPT diagram \citep{Stasinska:2008}. We apply an H$\alpha$ equivalent width (EW) cut of \ewha\ $>$ 3\AA\ to remove these retired galaxies \citet{Cid-Fernandes:2011}. In this work, we define star forming galaxies as those which are in the blue cloud in the CMD and which are below the theoretical maximum starburst line from \citet{Kewley:2001} on the BPT diagram.

We measure the SFR from the H$\alpha$ luminosity, L$_{H\alpha}$, calculated from the H$\alpha$ flux.  We use the SFR formula, Equation \ref{eq:sfr}, from \citet{Murphy:2011} which uses a Kroupa IMF, Solar metallicity, a constant SFR at an age of 100 Myr, and Case-B recombination.

@@

Since the stellar mass of a galaxy is not uniformly distributed within the galaxy, we normalize the SFR by the local stellar mass, giving us the specific star formation rate (sSFR). The local stellar masses used here is the stellar mass calculated by {\sc SPFIT}'s fit to the stellar continuum with each spaxel.

\begin{equation}
{\rm sSFR} = \frac{\rm SFR}{M^*}
\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Pair Sample}\label{sec:pair}
We build two different pair samples in this work; the inside-IFU sample which contains paired galaxies where both galaxies are contained within a single MaNGA IFU and the outside-IFU sample where a MaNGA target galaxies is coupled with another galaxies found outside of the MaNGA IFU.

@@

We use SDSS photometric objects to create a catalog of objects within the fields of view of the MaNGA IFUs. We extract their spectra from the MaNGA datacubes through a 1\arcsec\ aperture and fit the spectra with {\sc SPFIT} assuming the MaNGA target's redshift. The spectra is then sorted into different categories; ``good" galaxy spectra, broad-line AGN, foreground star, foreground/background galaxies, or poor S/N objects. The ``good" galaxy spectra are the objects whose spectra is well modeled by {\sc SPFIT} at the target galaxy's redshift, whether it is the target galaxy itself or a nearby companion galaxy. This means that the companion galaxy can be within $\pm$3000 km s$^{-1}$ of the MaNGA target. We found 6573 ``good" objects, 57 broad-line AGN, 836 foreground stars, 319 foreground/background galaxies, and 1546 objects with poor S/N.

From the 6573 galaxies with good spectra, 404 of the MaNGA IFUs have multiple objects within the IFU. We further restrict the sample by setting a relative velocity cut of $\Delta$v $<$ 500 km s$^{-1}$ to remove projected companions and we also require that the galaxies are classified as star forming as described in Section \ref{sec:analysis}. These requirements reduce our sample down to 124 star forming galaxies (SFGs).

Several of the target galaxies in our sample have multiple companions. From the whole set of 404 IFUs, there are 327 pairs, 67 triplets, 7 quadruplets, and 1 quintuplet.

We characterize the geometry of the inside-IFU galaxies with the r-band elliptical S\'ersic apertures and we characterize the total stellar mass of the galaxies with the stellar masses from \citet{Mendel:2014}. The Simard+11 and \citet{Mendel:2014} catalogs do not completely cover the MaNGA sample. Out of the 124 SFGs in the inside-IFU sample, 41 of the SFGs are covered by both \simard\ and \citet{Mendel:2014}.

@@

We supplement the inside-IFU sample with a set of pairs identified outside of the field of view of the MaNGA IFU. We select these outside-IFU pairs from the NSA catalog. We search for these external pairs by selecting objects with a projected separation from the MaNGA targets of r$_p$ $<$ 50 kpc using the MaNGA target's redshift. We further use a relative velocity cut of $\Delta$v $<$ 500 km s$^{-1}$ to remove projected galaxies from the selection.

From the NSA catalog's 641,409 galaxies, we find 509 galaxies which are paired to MaNGA targets. MaNGA targets which have both an inside-IFU and an outside-IFU pair are left to the inside-IFU sample. After restricting the sample to SFGs, we have 161 MaNGA targets with paired galaxies outside of the IFU.

We use the r-band elliptical Petrosian apertures from the NSA catalog to describe the geometry of the galaxies in the outside-IFU sample. We also use the stellar masses calculated from the r-band elliptical Petrosian apertures in the NSA catalog for the total stellar mass of the galaxies.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Control Sample}\label{sec:control}
We build a sample of isolated control galaxies from the remaining MaNGA galaxies by selecting the MaNGA target galaxies which have no spectroscopic companions within r$_p$ $<$ 50 kpc and $\Delta$v $<$ 500 km s$^{-1}$ either inside or outside of the IFU. This gives us a set of 1862 SFGs and 1598 SFGs which are covered by both \simard\ and \citet{Mendel:2014}.

The SFR in galaxies has been shown to be dependent on both stellar mass and redshift of the galaxies \citep{Noeske:2007}. This effect needs to be controlled for when comparing the paired and control galaxies to isolate the changes in the sSFR due to merger induced effects. For each galaxy pair we will build a subset of 20 control galaxies which are matched on stellar mass and redshift.

To control for stellar mass, we set a 0.1 dex stellar mass cut between a paired galaxy and the potential control galaxies. To restrict the redshift between the control and paired galaxies, we select control galaxies which are drawn from the same MaNGA subsample (e.g. Primary or Secondary) as the paired galaxy. The MaNGA sample has a tight ``banana" shaped distribution in stellar mass and redshift, seen in Figure \ref{fig:Mi-z}. By setting a 0.1 dex stellar mass cut and requiring that the galaxies be in the same MaNGA subsample, we effectively set a z $\le$ $\sim$0.025 redshift cut. We note that the MaNGA galaxies with stellar masses above \logm\ $\ge$ 11.0 have a wide redshift distribution. In this work, we focus on the galaxies below that mass range.

Setting the above requirements, most paired galaxies will find between 20$-$100 control galaxies. Since we want each paired galaxy to be treated in a similar manner, we will downselect the total number of acquired control galaxies down to 20 controls. We select the 20 controls using a random number generator. In the cases where a paired galaxy does not initially acquire 20 control galaxies we iteratively expand the stellar mass limit by 0.1 dex until at least 20 control galaxies are found.

A small set of paired required extra iterations to find 20 controls; 8 paired galaxies need an extra iteration, 1 paired galaxy need two extra iterations, and 1 paired galaxy needed three iterations.

The control galaxies have their radial profiles of sSFR made with both the r-band elliptical Petrosian apertures from the NSA catalog and the r-band elliptical S\'ersic apertures from \simard. We use the geometries and stellar masses from \simard\ and \citet{Mendel:2014} when the controls are being matched to pairs in the inside-IFU sample and we use the geometries and stellar masses from the NSA catalog when the controls are being matched to pairs in the outside-IFU sample.

@@

In order to spatially characterize the star formation in the paired galaxies we build radial profiles of sSFR. As mentioned in Sections \ref{sec:pair} and \ref{sec:control} we use the r-band elliptical Petrosian apertures from the NSA catalog and the r-band S\'ersic apertures from \simard\ to define the geometries of the galaxies. We calculate the inclination angle, {\it i}, of the galaxies using the major-to-minor axis ratios from the elliptical apertures;

@@

Where $q$ is the intrinsic oblateness. We use the empirically determined oblateness of $q = 0.13$, from \citet{Giovanelli:1994}.

The inclination angle, along with the galaxy's position angle, is used to deproject the geometries of the galaxies. The distance to each of the spaxels from the center of the MaNGA target is calculated using the 50\% half light radius to scale the size of the galaxy. The r-band elliptical S\'ersic apertures from \simard\ are used for the inside-IFU sample and their controls and the r-band elliptical Petrosian apertures from the NSA catalog are used for the outside-IFU sample and their controls.

The spaxels are binned into radius increments of 0.1 \reff\ from 0.0$-$2.5 \reff. Within each radius bin we take the median of the specific star formation rate. The profiles' errors are the standard error of the mean of the data within each radius bin. We create one of these individual averaged radial profiles for every MaNGA galaxy.
